---
title: "Exploratory analysis"
author: "Ana Abad"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(gtsummary)
library(ggplot2)
library(ggpubr)

```


```{r}
load("ATP_final.RData")
```
```{r}

```


## 1. Summary
```{r}
summary(datos_combinados)
```

## 2. Variable characteristics


### games


```{r}
table(datos_combinados$games, useNA = "always")
```


### tourney_name

```{r}
n_distinct(datos_combinados$tourney_name)
n_distinct(datos_combinados$tourney_id)
```

### Surface
```{r}
unique(datos_combinados$surface)
table(datos_combinados$surface)
```
### tourney_level and level

```{r}

unique(datos_combinados$tourney_level)
datos_combinados$tourney_level <- droplevels(datos_combinados$tourney_level)
table(datos_combinados$tourney_level)

```

### round

```{r}
datos_combinados <- datos_combinados %>%
  mutate(round = case_when(
    round %in% c("Qualifier") ~ "Qualifying",
    round %in% c("Preliminary") ~ "Preliminary",
    round %in% c("QuarterFinal", "SemiFinal", "Final") ~ "Final",
    TRUE ~ NA_character_
  ))
table(datos_combinados$round)
```
### hands 

```{r}

unique(datos_combinados$winner_hand)
unique(datos_combinados$loser_hand)

```

### Age
```{r}
summary(datos_combinados$winner_age)
summary(datos_combinados$loser_age)
is.integer(datos_combinados$winner_rank)
summary(datos_combinados$dif_age)
```


### Ranking


```{r}

missing_data_summary <- datos_combinados %>%
  summarize(
    winner_rank_missing = sum(is.na(winner_rank)),
    loser_rank_missing = sum(is.na(loser_rank)),
    .by = year
  )


missing_data_long <- missing_data_summary %>%
  pivot_longer(cols = c(winner_rank_missing, loser_rank_missing), 
               names_to = "rank_type", 
               values_to = "missing_count")

missrankbyyear <- ggplot(missing_data_long, aes(x = year, y = missing_count, color = rank_type, group = rank_type)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c("winner_rank_missing" = "#000078", "loser_rank_missing" = "#73edff"),
    labels = c("Winner Rank Missing", "Loser Rank Missing")
  ) +
  labs(
    title = "Missing Values in Player Ranks Over Time",
    x = "Year",
    y = "Number of Missing Values",
    color = "Rank Type"
  ) +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(
    text = element_text(size = 12),
    legend.position = "top"
  )




ggsave(filename = "Rankbyyear.tif", plot = missrankbyyear, device = "tiff",  dpi = 300)

```

### winner/loser_entry

```{r}
unique(datos_combinados$winner_entry)
unique(datos_combinados$loser_entry)
datos_combinados <- datos_combinados %>%
  mutate(
    winner_entry = case_when(
      # Lucky Loser
      winner_entry %in% c("LL", "Lucky Loser") ~ "Lucky Loser",

      # Protected Ranking
      winner_entry %in% c("PR", "Protected Ranking") ~ "Protected Ranking",

      # Qualifier
      winner_entry %in% c("Q", "Qualifier") ~ "Qualifier",

      # Special Exempt
      winner_entry %in% c("SE", "S", "Special Exempt") ~ "Special Exempt",

      # Wild Card
      winner_entry %in% c("WC", "W", "Wild Card") ~ "Wild Card",

      # Alternate (todas las variantes encontradas)
      winner_entry %in% c("ALT", "AL", "Alt", "Alternate",
                          "2) A", "5) A", "A") ~ "Alternate",

      # Sin código → Direct Entry
      is.na(winner_entry) | winner_entry == "" ~ "Direct Entry",

      # En cualquier otro caso, conservar (p. ej. ya limpio)
      TRUE ~ winner_entry
    ),

    loser_entry = case_when(
      # Lucky Loser
      loser_entry %in% c("LL", "Lucky Loser") ~ "Lucky Loser",

      # Protected Ranking
      loser_entry %in% c("PR", "Protected Ranking") ~ "Protected Ranking",

      # Qualifier
      loser_entry %in% c("Q", "Qualifier") ~ "Qualifier",

      # Special Exempt
      loser_entry %in% c("SE", "S", "Special Exempt") ~ "Special Exempt",

      # Wild Card
      loser_entry %in% c("WC", "W", "Wild Card") ~ "Wild Card",

      # Alternate (todas las variantes encontradas)
      loser_entry %in% c("ALT", "AL", "Alt", "Alternate",
                         "2) A", "5) A", "A") ~ "Alternate",

      # Sin código → Direct Entry
      is.na(loser_entry) | loser_entry == "" ~ "Direct Entry",

      TRUE ~ loser_entry
    )
  )
table(datos_combinados$loser_entry)
table(datos_combinados$winner_entry)


datos_combinados <- datos_combinados %>%
  mutate(
    # Lucky Loser flags
    LLW = if_else(winner_entry == "Lucky Loser", 1, 0),
    LLL = if_else(loser_entry  == "Lucky Loser", 1, 0),

    # Protected Ranking flags
    PRW = if_else(winner_entry == "Protected Ranking", 1, 0),
    PRL = if_else(loser_entry  == "Protected Ranking", 1, 0),

    # Summary variables (Yes/No)
    LL = if_else(LLW == 1 | LLL == 1, "Yes", "No"),
    PR = if_else(PRW == 1 | PRL == 1, "Yes", "No")
  )

table(datos_combinados$LL)
table(datos_combinados$PR)


primer_pr <- datos_combinados %>%
  filter(
    winner_entry == "Protected Ranking" |
      loser_entry == "Protected Ranking"
  ) %>%
  summarise(
    primer_anio_PR = min(year, na.rm = TRUE),
    ultimo_anio_PR = max(year, na.rm = TRUE),
    n_partidos_PR  = n()
  )

primer_pr

datos_PR_periodo <- datos_combinados %>%
  filter(year >= 2000)



save(datos_PR_periodo, file = "datos_combinados.RData")

```



### Summary 


```{r}

tbl_summary(
  data    = datos_PR_periodo,
  include = c(
    surface,
    tourney_level,
    winner_hand,
    loser_hand,
    dif_age,
    winner_rank,
    loser_rank,
    dif_rank,
    round,
    best_of,
    games,
    winner_entry,
    loser_entry
  )
)
table(datos_PR_periodo$PR)
datos_PR_periodo$PR_both <- with(datos_PR_periodo,
  ifelse(PRW == 1 & PRL == 1, "Both PR",
  ifelse(PRW == 1 | PRL == 1, "One PR", "No PR"))
)
table(datos_PR_periodo$PR_both)

```
```{r}

tbl_summary(
  data    = datos_PR_periodo,
  include = c(
    surface,
    tourney_level,
    winner_hand,
    loser_hand,
    dif_age,
    winner_rank,
    loser_rank,
    dif_rank,
    round,
    best_of,
    games,
    winner_entry,
    loser_entry
  ),
  by = PR
)

```


## 3. Distributions



```{r}

plot_histogram(datos_PR_periodo)
plot_bar(datos_PR_periodo)


```

### Normality in continuous variables

```{r}
ggqqplot(datos_PR_periodo$dif_age, title ="dif_age")
ks.test(datos_PR_periodo$dif_age, "pnorm", mean(datos_PR_periodo$dif_age, na.rm = TRUE), sd(datos_PR_periodo$dif_age, na.rm = TRUE))
```
dif_age does not follow a normal distribution. 


```{r}
ggqqplot(datos_PR_periodo$dif_rank, title = "dif_rank")
ks.test(datos_PR_periodo$dif_rank, "pnorm", mean(datos_PR_periodo$dif_rank, na.rm = TRUE), sd(datos_PR_periodo$dif_rank, na.rm = TRUE))
```
dif_rank does not follow a normal distribution.

```{r}
ggqqplot(datos_PR_periodo$winner_age, title = "winner_age")
ks.test(datos_PR_periodo$winner_age, "pnorm", mean(datos_PR_periodo$winner_age, na.rm = TRUE), sd(datos_PR_periodo$winner_age, na.rm = TRUE))
```
winner_age does not follow a normal distribution.


```{r}
ggqqplot(datos_PR_periodo$loser_age, title = "loser_age")
ks.test(datos_PR_periodo$loser_age, "pnorm", mean(datos_PR_periodo$loser_age, na.rm = TRUE), sd(datos_PR_periodo$loser_age, na.rm = TRUE))
```
loser_age does not follow a normal distribution

```{r}
ggqqplot(datos_PR_periodo$winner_rank, title = "winner_rank")
ks.test(datos_PR_periodo$winner_rank, "pnorm", mean(datos_PR_periodo$winner_rank, na.rm = TRUE), sd(datos_PR_periodo$winner_rank, na.rm = TRUE))
```
winner_rank does not follow a normal distribution

```{r}
ggqqplot(datos_PR_periodo$loser_rank, title = "loser_rank")
ks.test(datos_PR_periodo$loser_rank, "pnorm", mean(datos_PR_periodo$loser_rank, na.rm = TRUE), sd(datos_PR_periodo$loser_rank, na.rm = TRUE))
```

loser_rank does not follow a normal distribution
```{r}
ggqqplot(datos_PR_periodo$best_of, title = "best_of")
ks.test(ddatos_PR_periodo$best_of, "pnorm", mean(datos_PR_periodo$best_of, na.rm = TRUE), sd(datos_PR_periodo$best_of, na.rm = TRUE))
```
best_of does not follow a normal distribution


```{r}
ggqqplot(datos_PR_periodo$games, title = "games")

ks.test(datos_PR_periodo$games, "pnorm", mean(datos_PR_periodo$games, na.rm = TRUE), sd(datos_PR_periodo$games, na.rm = TRUE))

```

games does not follow a normal distribution.

## 5. Missing data


```{r}
missing_per_column <- colSums(is.na(datos_PR_periodo))

missing_percentage <- missing_per_column * 100 / nrow(datos_PR_periodo) 
missing_percentage


missing_data <- data.frame(
  variable = names(datos_PR_periodo),
  missing_count = colSums(is.na(datos_PR_periodo)),
  total = nrow(datos_PR_periodo)
)

missing_data$percentage <- missing_percentage

Missingdata <- ggplot(missing_data, aes(x = reorder(variable, -percentage), y = percentage)) +
  geom_bar(stat = "identity", fill = "#73edff") +
  labs(title = "Missing data in each variable", 
       x = "Variable", 
       y = "Missing data (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90))

Missingdata

ggsave(filename = "Missingdata.tif", plot = Missingdata, device = "tiff",  dpi = 300)
datos_combinados<- datos_PR_periodo

```


