---
title: "Descriptive analysis"
author: "Ana Abad"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(compareGroups)
library(patchwork)
library(ggdist)
library(gghalves)
library(epiR)
library(SmartEDA)



```



```{r}
load("ATP_final.RData")
```


```{r}
datos_combinados <- datos_combinados %>%
  mutate(round = case_when(
    grepl("^R[0-9]+$", round) ~ "Preliminary", 
    round %in% c("RR", "BR") ~ "Preliminary",  
    TRUE ~ round
  ))

datos_combinados <- datos_combinados %>%
  mutate(round = case_when(
    round %in% c("Q1", "Q2", "Q3", "Q4") ~ "Qualifier",  
    TRUE ~ round
  ))


datos_combinados <- datos_combinados %>%
  mutate(round = recode(round, 
                        "QF" = "QuarterFinal",
                        "SF" = "SemiFinal",
                        "F" = "Final"))

```




# 1. Mean

```{r}
round(sapply((datos_combinados %>% dplyr::select(where(is.numeric))), mean, na.rm = TRUE), 4)
```


# 2. Percentages: LL and PR

Is not useful, because LL/PR includes matches in which at least one of the players has the condition, but does not include information when the other player is also a LL/PR. Also, since the entry is for the whole tournament, it doesn't reflect the data correctly when using matches as a measurement unit.

```{r}
table(datos_combinados$PR)


table(datos_combinados$PR)["Yes"]*100/
  (table(datos_combinados$PR)["No"]+table(datos_combinados$PR)["Yes"])
  
```

```{r}
aux <- c("tourney_level", "year", "surface", "round", "games",
         "winner_hand", "loser_hand", "winner_age", "loser_age", "dif_age",
         "winner_rank", "loser_rank", "dif_rank", "match_outcome",
         "winner_entry", "loser_entry", "PR", "PRW", "PRL")
str(datos_combinados)
atp_eda <- datos_combinados[, aux]
rm(aux)

ExpReport(atp_eda, Template = NULL, Target = NULL, label = NULL, theme = "Default", op_file = "1 - Exploratory Data Analysis.html", op_dir = getwd(), sc = NULL, sn = NULL, Rc = NULL)
```
## Prevalence SE by year
```{r}
special_entries <- c("Alternate", "Lucky Loser", "Protected Ranking", "Qualifier", "Special Exempt", "Wild Card")

total_entries_per_year <- datos_combinados %>%
  filter(!is.na(year)) %>%
  count(year) %>%
  rename(total = n)

special_by_year <- datos_combinados %>%
  filter(winner_entry %in% special_entries | loser_entry %in% special_entries) %>%
  mutate(
    year = as.numeric(format(tourney_start_date, "%Y")),
    winner_entry = ifelse(winner_entry %in% special_entries, winner_entry, NA),
    loser_entry = ifelse(loser_entry %in% special_entries, loser_entry, NA)
  ) %>%
  select(year, winner_entry, loser_entry) %>%
  pivot_longer(cols = c(winner_entry, loser_entry), names_to = "role", values_to = "entry_type") %>%
  filter(!is.na(entry_type)) %>%
  count(year, entry_type) %>%
  left_join(total_entries_per_year, by = "year") %>%
  mutate(relative_frequency = n / total)

all_special <- special_by_year %>%
  group_by(year) %>%
  summarise(n = sum(n), total = first(total)) %>%
  mutate(entry_type = "All Special Entries", relative_frequency = n / total)

combined_fr <- bind_rows(special_by_year, all_special)

ggplot(combined_fr, aes(x = year, y = relative_frequency, color = entry_type)) +
  geom_line(aes(linetype = entry_type == "All Special Entries"), size = 1) +
  scale_linetype_manual(values = c("solid", "dashed"), guide = "none") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), breaks = seq(0, 1, by = 0.05)) +
  scale_x_continuous(breaks = seq(min(combined_fr$year), max(combined_fr$year), by = 5)) +
  labs(
    x = "Year",
    y = "Relative Frequency",
    color = "Entry Type"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

## Prevalence PR by year

```{r}

PR_summary <- datos_combinados %>%
  group_by(year) %>%
  summarise(
    PR_yes = sum(PR == "Yes"),
    PR_total = n(),
    PR_ratio = PR_yes / PR_total
  )

# 95 % CI

PR_summary <- PR_summary %>%
  rowwise() %>%
  mutate(
    conf_int = list(binom.test(PR_yes, PR_total)$conf.int),
    lower_ci = conf_int[[1]],
    upper_ci = conf_int[[2]]
  ) %>%
  ungroup()


```




```{r}
PR_summary$year <- as.numeric(PR_summary$year)

PRI <- ggplot(PR_summary, aes(x = year, y = PR_ratio)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
  labs(title = "Presence of Protected Ranking by Year", 
       x = "Year", 
       y = "Presence (PR = Yes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(limits = c(0, 0.03)) +
  scale_x_continuous(breaks = seq(min(PR_summary$year), max(PR_summary$year), by = 5))

PRI


```

## Prevalence PR by level

```{r}

PR_summary_l <- datos_combinados %>%
  group_by(tourney_level) %>%
  summarise(
    PR_yes = sum(PR == "Yes"),
    PR_total = n(),
    PR_ratio = PR_yes / PR_total
  )

# 95 % CI

PR_summary_l <- PR_summary_l %>%
  rowwise() %>%
  mutate(
    conf_int = list(binom.test(PR_yes, PR_total)$conf.int),
    lower_ci = conf_int[[1]],
    upper_ci = conf_int[[2]]
  ) %>%
  ungroup()


```


```{r}
PRlevel <- ggplot(PR_summary_l, aes(x = tourney_level, y = PR_ratio)) +
  geom_col() +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
  labs(title = "Prevalence of PR by level with 95% Confidence Interval", 
       x = "Year", 
       y = "Prevalence (PR = Yes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
   scale_y_continuous(limits = c(0, 0.01)) 

PRlevel

```


## Prevalence PR by age

```{r}

PR_summary_a <- datos_combinados %>%
  group_by(dif_age) %>%
  summarise(
    PR_yes = sum(PR == "Yes"),
    PR_total = n(),
    PR_ratio = PR_yes / PR_total
  )

# 95 % CI

PR_summary_a <- PR_summary_a %>%
  rowwise() %>%
  mutate(
    conf_int = list(binom.test(PR_yes, PR_total)$conf.int),
    lower_ci = conf_int[[1]],
    upper_ci = conf_int[[2]]
  ) %>%
  ungroup()


```


```{r}
PRage <- ggplot(PR_summary_a, aes(x = dif_age, y = PR_ratio)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
  labs(title = "Prevalence of Lucky loser by age with 95% Confidence Interval", 
       x = "Age difference", 
       y = "Prevalence (PR = Yes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
   scale_y_continuous(limits = c(0, 0.02)) 

PRage

```


## PR by hand

### Winner
```{r}

PR_summary_hw <- datos_combinados %>%
  group_by(winner_hand) %>%
  summarise(
    PR_yes = sum(PR == "Yes"),
    PR_total = n(),
    PR_ratio = PR_yes / PR_total
  )

# 95 % CI

PR_summary_hw <- PR_summary_hw %>%
  rowwise() %>%
  mutate(
    conf_int = list(binom.test(PR_yes, PR_total)$conf.int),
    lower_ci = conf_int[[1]],
    upper_ci = conf_int[[2]]
  ) %>%
  ungroup()


```


```{r}
PRwh <- ggplot(PR_summary_hw, aes(x = winner_hand, y = PR_ratio)) +
  geom_col() +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
  labs(title = "Prevalence of Lucky loser by age with 95% Confidence Interval", 
       x = "Winner Hand", 
       y = "Prevalence (PR = Yes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
   scale_y_continuous(limits = c(0, 0.05)) 

PRwh

```
### Loser


```{r}

PR_summary_hl <- datos_combinados %>%
  group_by(loser_hand) %>%
  summarise(
    PR_yes = sum(PR == "Yes"),
    PR_total = n(),
    PR_ratio = PR_yes / PR_total
  )

# 95 % CI

PR_summary_hl <- PR_summary_hl %>%
  rowwise() %>%
  mutate(
    conf_int = list(binom.test(PR_yes, PR_total)$conf.int),
    lower_ci = conf_int[[1]],
    upper_ci = conf_int[[2]]
  ) %>%
  ungroup()


```


```{r}
PRlh <- ggplot(PR_summary_hl, aes(x = loser_hand, y = PR_ratio)) +
  geom_col() +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
  labs(title = "Prevalence of Lucky loser by age with 95% Confidence Interval", 
       x = "Loser Hand", 
       y = "Prevalence (PR = Yes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
   scale_y_continuous(limits = c(0, 0.05)) 

PRlh

```

## PR by surface

```{r}

PR_summary_sur <- datos_combinados %>%
  group_by(surface) %>%
  summarise(
    PR_yes = sum(PR == "Yes"),
    PR_total = n(),
    PR_ratio = PR_yes / PR_total
  )

# 95 % CI

PR_summary_sur <- PR_summary_sur %>%
  rowwise() %>%
  mutate(
    conf_int = list(binom.test(PR_yes, PR_total)$conf.int),
    lower_ci = conf_int[[1]],
    upper_ci = conf_int[[2]]
  ) %>%
  ungroup()

PR_summary_sur
```


```{r}
PRsur <- ggplot(PR_summary_sur, aes(x = surface, y = PR_ratio)) +
  geom_col() +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
  labs(title = "Prevalence of Lucky loser by age with 95% Confidence Interval", 
       x = "Surface", 
       y = "Prevalence (PR = Yes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
   scale_y_continuous(limits = c(0, 0.05)) 

PRsur

```


## PR by country

### Winner
```{r}

PR_summary_cw <- datos_combinados %>%
  group_by(winner_ioc) %>%
  summarise(
    PR_yes = sum(PR == "Yes"),
    PR_total = n(),
    PR_ratio = PR_yes / PR_total
  )

# 95 % CI

PR_summary_cw <- PR_summary_cw %>%
  rowwise() %>%
  mutate(
    conf_int = list(binom.test(PR_yes, PR_total)$conf.int),
    lower_ci = conf_int[[1]],
    upper_ci = conf_int[[2]]
  ) %>%
  ungroup()


```


```{r}
PRcw <- ggplot(PR_summary_cw, aes(x = winner_ioc, y = PR_ratio)) +
  geom_col() +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
  labs(title = "Prevalence of Lucky loser by age with 95% Confidence Interval", 
       x = "Winner ioc", 
       y = "Prevalence (PR = Yes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
   scale_y_continuous(limits = c(0, 0.05)) 

PRcw

```
### Loser


```{r}

PR_summary_cl <- datos_combinados %>%
  group_by(loser_ioc) %>%
  summarise(
    PR_yes = sum(PR == "Yes"),
    PR_total = n(),
    PR_ratio = PR_yes / PR_total
  )

# 95 % CI

PR_summary_cl <- PR_summary_cl %>%
  rowwise() %>%
  mutate(
    conf_int = list(binom.test(PR_yes, PR_total)$conf.int),
    lower_ci = conf_int[[1]],
    upper_ci = conf_int[[2]]
  ) %>%
  ungroup()


```


```{r}
PRlc <- ggplot(PR_summary_cl, aes(x = loser_ioc, y = PR_ratio)) +
  geom_col() +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
  labs(title = "Prevalence of Lucky loser by age with 95% Confidence Interval", 
       x = "Loser ioc", 
       y = "Prevalence (PR = Yes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90)) +
   scale_y_continuous(limits = c(0, 0.05)) 

PRlc

```


# 3. Percentaje per tourney
```{r}

# Create a new dataframe that counts unique PR per tournament
unique_pr_per_tournament <- datos_combinados %>%
  # Create a temporary dataframe to identify PR players in each match
  dplyr::select(tourney_id, winner_id, loser_id, PRW, PRL) %>%
  pivot_longer(cols = c(winner_id, loser_id), names_to = "player_type", values_to = "player_id") %>%
  # Keep only matches where players are PR
  filter((player_type == "winner_id" & PRW == 1) | (player_type == "loser_id" & PRL == 1)) %>%
  # Group by tournament and count unique PR players
  group_by(tourney_id) %>%
  summarise(unique_PR_count = n_distinct(player_id)) %>%
  ungroup()



sum(unique_pr_per_tournament$unique_PR_count)
n_distinct(unique_pr_per_tournament$tourney_id)
n_distinct(datos_combinados$tourney_id)


Tournaments_with_PR <- n_distinct(unique_pr_per_tournament$tourney_id)*100/n_distinct(datos_combinados$tourney_id)
Tournaments_with_PR

binom.test(sum(unique_pr_per_tournament$unique_PR_count), n_distinct(datos_combinados$tourney_id), conf.level = 0.95)


```


# 4. Winer/loser new variable

```{r}

datos_combinados <- datos_combinados %>%
  mutate(
    PR_wins = case_when(
      winner_entry == "PR" ~ 1,
      loser_entry == "PR" ~ 0,
      TRUE ~ NA_real_
    ),
    PR_loses = case_when(
      loser_entry == "PR" ~ 1,
      winner_entry == "PR" ~ 0,
      TRUE ~ NA_real_
    )
  )


```

## Comparing groups

```{r}
compareGroups(PRW ~ ., method = NA, data = datos_combinados)
compareGroups(PRL ~ ., method = NA, data = datos_combinados)

```



```{r}

descrTable(PR_wins ~ surface + tourney_level + winner_id + winner_entry + winner_hand 
           + winner_age + loser_id +  loser_entry + loser_hand + loser_age + best_of 
           + round  + winner_rank + loser_rank + games + dif_age + dif_rank, data = datos_combinados, byrow = TRUE, show.p.overall = FALSE, digits = 2, method = c(mean_age=NA), alpha=0.05)

```


# 5. Special entries matches 

```{r}
categorize_match <- function(winner_type, loser_type) {
  if (winner_type == "Lucky Loser" & loser_type == "Lucky Loser") {
    return("LL & LL")
  } else if ((winner_type == "Lucky Loser" & loser_type == "Protected Ranking") |
             (winner_type == "Protected Ranking" & loser_type == "Lucky Loser")) {
    return("LL & PR")
  } else if (winner_type == "Protected Ranking" & loser_type == "Protected Ranking") {
    return("PR & PR")
  } else {
    return(NA)  
  }
}


datos_combinados <- datos_combinados %>%
  mutate(match_type = mapply(categorize_match, winner_entry, loser_entry))

# Tabla resumen
table_summary <- datos_combinados %>%
  group_by(match_type) %>%
  summarize(nmatches = n(), .groups = "drop") %>%
  pivot_wider(names_from = match_type, values_from = nmatches, 
              values_fill = list(nmatches = 0)) %>%
  rename(
    `Lucky Loser` = `LL & LL`, 
    `Protected Ranking` = `PR & PR`, 
    `LL & PR` = `LL & PR`
  )

# Mostrar resultado
table_summary

```


# 6. Summary 



```{r}

tbl_summary(data = datos_PR_periodo, by = PR, include = c(year, surface, tourney_level, winner_hand, loser_hand, dif_age, winner_rank, loser_rank, best_of))


tbl_summary(data = datos_combinados, by = LL, include = c(year, surface, tourney_level, winner_hand, loser_hand, dif_age, winner_rank, loser_rank))



```

# 7. Graphics


# Raincloud plot 

## dig_age

```{r, eval= FALSE}


ggplot(datos_combinados, aes(x = LL, y = dif_age)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    ## draw jitter on the left
    side = "l", 
    ## control range of jitter
    range_scale = .4, 
    ## add some transparency
    alpha = .2
  ) +
  coord_cartesian(xlim = c(1.2, 2.9), clip = "off")





ggplot(datos_combinados, aes(x = PR, y = dif_age)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    ## draw jitter on the left
    side = "l", 
    ## control range of jitter
    range_scale = .4, 
    ## add some transparency
    alpha = .2
  ) +
  coord_cartesian(xlim = c(1.2, 2.9), clip = "off")

```

```{r}


plot_LL <- ggplot(datos_combinados, aes(x = LL, y = dif_age)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  gghalves::geom_half_point(
    side = "l", 
    range_scale = .4, 
    alpha = .2
  ) +
  coord_cartesian(xlim = c(1.2, 2.9), clip = "off")


plot_PR <- ggplot(datos_combinados, aes(x = PR, y = dif_age)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  gghalves::geom_half_point(
    side = "l", 
    range_scale = .4, 
    alpha = .2
  ) +
  coord_cartesian(xlim = c(1.2, 2.9), clip = "off")


combined_plot <- plot_LL + plot_PR


ggsave("combined_plot.tiff", plot = combined_plot, width = 12, height = 6, dpi = 300, device = "tiff")

```


## Rank diference


```{r, eval= FALSE}

# Create the LL plot for dif_rank
plot_LL_dif_rank <- ggplot(datos_combinados, aes(x = LL, y = dif_rank)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  gghalves::geom_half_point(
    side = "l", 
    range_scale = .4, 
    alpha = .2
  ) +
  coord_cartesian(xlim = c(1.2, 2.9), clip = "off")

# Create the PR plot for dif_rank
plot_PR_dif_rank <- ggplot(datos_combinados, aes(x = PR, y = dif_rank)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  gghalves::geom_half_point(
    side = "l", 
    range_scale = .4, 
    alpha = .2
  ) +
  coord_cartesian(xlim = c(1.2, 2.9), clip = "off")

# Combine both plots side by side
combined_dif_rank_plot <- plot_LL_dif_rank + plot_PR_dif_rank

# Save the combined plot as a TIFF file
ggsave("combined_dif_rank_plot.tiff", plot = combined_dif_rank_plot, width = 12, height = 6, dpi = 300, device = "tiff")


```






## Age AND Rank differnce

```{r, eval= FALSE}
ggplot(datos_combinados, aes(x = dif_rank, y = dif_age, color = factor(LL))) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  gghalves::geom_half_point(
    side = "l", 
    range_scale = .4, 
    alpha = .2
  ) +
  facet_wrap(~ LL, scales = "free", labeller = labeller(LL = c(`0` = "No", `1` = "Yes"))) +
  labs(
    x = "Difference in Rank",
    y = "Difference in Age",
    color = "Match Outcome"
  ) +
  coord_cartesian(clip = "off") +
  theme_minimal()

```



## Surface




```{r, eval= FALSE}

ggplot(datos_combinados, aes(x = LL, y = surface)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    ## draw jitter on the left
    side = "l", 
    ## control range of jitter
    range_scale = .4, 
    ## add some transparency
    alpha = .2
  ) +
  coord_cartesian(xlim = c(1.2, 2.9), clip = "off")





ggplot(datos_combinados, aes(x = PR, y = surface)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    ## draw jitter on the left
    side = "l", 
    ## control range of jitter
    range_scale = .4, 
    ## add some transparency
    alpha = .2
  ) +
  coord_cartesian(xlim = c(1.2, 2.9), clip = "off")

```

```{r}

plot_LL_surface <- ggplot(datos_combinados, aes(x = LL, y = surface)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    ## draw jitter on the left
    side = "l", 
    ## control range of jitter
    range_scale = .4, 
    ## add some transparency
    alpha = .2
  ) +
  coord_cartesian(xlim = c(1.2, 2.9), clip = "off")





plot_PR_surface <- ggplot(datos_combinados, aes(x = PR, y = surface)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    ## draw jitter on the left
    side = "l", 
    ## control range of jitter
    range_scale = .4, 
    ## add some transparency
    alpha = .2
  ) +
  coord_cartesian(xlim = c(1.2, 2.9), clip = "off")



# Combine both plots using patchwork
combined_surface_plot <- plot_LL_surface + plot_PR_surface

# Save the combined plot as a TIFF file
ggsave("combined_surface_plot.tiff", plot = combined_surface_plot, width = 12, height = 6, dpi = 300, device = "tiff")

```






## Tourney level



```{r, eval= FALSE}

ggplot(datos_combinados, aes(x = LL, y = tourney_level)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    ## draw jitter on the left
    side = "l", 
    ## control range of jitter
    range_scale = .4, 
    ## add some transparency
    alpha = .2
  ) +
  coord_cartesian(xlim = c(1.2, 2.9), clip = "off")





ggplot(datos_combinados, aes(x = PR, y = tourney_level)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) + 
  geom_boxplot(
    width = .15, 
    outlier.shape = NA
  ) +
  ## add justified jitter from the {gghalves} package
  gghalves::geom_half_point(
    ## draw jitter on the left
    side = "l", 
    ## control range of jitter
    range_scale = .4, 
    ## add some transparency
    alpha = .2
  ) +
  coord_cartesian(xlim = c(1.2, 2.9), clip = "off")

```


# 8. Epi2by2


## PR 


```{r}
torneos <- datos_combinados %>%
  group_by(tourney_id) %>%
  summarise(
    PR_torneo = if_else(any(PR == "Yes"), "Yes", "No"),
    surface = first(surface),
    tourney_level = first(tourney_level),
    best_of = first(best_of),
    .groups = "drop"
  )

table(torneos$PR_torneo)


tabla_PR_torneo <- table(torneos$PR_torneo)

n_torneos_total    <- sum(tabla_PR_torneo)
n_torneos_con_PR   <- tabla_PR_torneo["Yes"]

IP_global <- n_torneos_con_PR / n_torneos_total


IP_global_1000 <- IP_global * 1000
IP_global_1000

IC_global <- binom.test(n_torneos_con_PR, n_torneos_total, conf.level = 0.95)
IC_global


IC_global_1000 <- IC_global$conf.int * 1000
IC_global_1000

#Surface
(tabla_surface_PR_t <- table(torneos$surface, torneos$PR_torneo))
Valores_Grass_Clay_PR_t <- c(
  tabla_surface_PR_t["Clay",  "Yes"],
  tabla_surface_PR_t["Clay",  "No"],
  tabla_surface_PR_t["Grass", "Yes"],
  tabla_surface_PR_t["Grass", "No"]
)

epi.2by2(
  dat        = Valores_Grass_Clay_PR_t,
  method     = "cohort.count",
  digits     = 2,
  conf.level = 0.95,
  units      = 1000,       # IP y RD por 1000 torneos
  interpret  = FALSE,
  outcome    = "as.columns"
)

Valores_Grass_Hard_PR_t <- c(
  tabla_surface_PR_t["Hard",  "Yes"],
  tabla_surface_PR_t["Hard",  "No"],
  tabla_surface_PR_t["Grass", "Yes"],
  tabla_surface_PR_t["Grass", "No"]
)

epi.2by2(
  dat        = Valores_Grass_Hard_PR_t,
  method     = "cohort.count",
  digits     = 2,
  conf.level = 0.95,
  units      = 1000,
  interpret  = FALSE,
  outcome    = "as.columns"
)

Valores_Grass_Carpet_PR_t <- c(
  tabla_surface_PR_t["Carpet", "Yes"],
  tabla_surface_PR_t["Carpet", "No"],
  tabla_surface_PR_t["Grass",  "Yes"],
  tabla_surface_PR_t["Grass",  "No"]
)

epi.2by2(
  dat        = Valores_Grass_Carpet_PR_t,
  method     = "cohort.count",
  digits     = 2,
  conf.level = 0.95,
  units      = 1000,
  interpret  = FALSE,
  outcome    = "as.columns"
)

#Tourney level

(tabla_tourney_level_PR_t <- table(torneos$tourney_level, torneos$PR_torneo))

Valores_M_A_PR_t <- c(
  tabla_tourney_level_PR_t["250 or 500", "Yes"],
  tabla_tourney_level_PR_t["250 or 500", "No"],
  tabla_tourney_level_PR_t["Masters",     "Yes"],
  tabla_tourney_level_PR_t["Masters",     "No"]
)

epi.2by2(
  dat        = Valores_M_A_PR_t,
  method     = "cohort.count",
  digits     = 2,
  conf.level = 0.95,
  units      = 1000,
  interpret  = FALSE,
  outcome    = "as.columns"
)

Valores_M_G_PR_t <- c(
  tabla_tourney_level_PR_t["Grand Slams", "Yes"],
  tabla_tourney_level_PR_t["Grand Slams", "No"],
  tabla_tourney_level_PR_t["Masters",     "Yes"],
  tabla_tourney_level_PR_t["Masters",     "No"]
)

epi.2by2(
  dat        = Valores_M_G_PR_t,
  method     = "cohort.count",
  digits     = 2,
  conf.level = 0.95,
  units      = 1000,
  interpret  = FALSE,
  outcome    = "as.columns"
)

Valores_M_F_PR_t <- c(
  tabla_tourney_level_PR_t["Tour Finals", "Yes"],
  tabla_tourney_level_PR_t["Tour Finals", "No"],
  tabla_tourney_level_PR_t["Masters",     "Yes"],
  tabla_tourney_level_PR_t["Masters",     "No"]
)

epi.2by2(
  dat        = Valores_M_F_PR_t,
  method     = "cohort.count",
  digits     = 2,
  conf.level = 0.95,
  units      = 1000,
  interpret  = FALSE,
  outcome    = "as.columns"
)

```

# 9. Proability of winning 

```{r}

# LL vs PR

matches_LL_PR <- datos_combinados %>%
  filter(
    LL == "Yes" &  # The match involves a lucky loser (LL)
    (winner_entry == "PR" | loser_entry == "PR")  # One of the players is a protected ranking player (PR)
  )

# Number of LL wins against PR / total number of matches involving LL against PR
prop.test(sum(matches_LL_PR$LLW == 1), nrow(matches_LL_PR), conf.level = 0.95)


# LL vs LL

matches_LL_vs_LL <-   datos_combinados %>% filter(
    LL == "Yes" &  (winner_entry == "LL" & loser_entry == "LL"))

prop.test(sum(matches_LL_vs_LL$LLW == 1), nrow(matches_LL_vs_LL), conf.level = 0.95)


# LL vs Direct entry

matches_LL_vs_DE <- datos_combinados[datos_combinados$LL == "Yes" & 
                                     (datos_combinados$winner_entry %in% c("Q", "WC", "ITF") | 
                                      datos_combinados$loser_entry %in% c("Q", "WC", "ITF")), ]


prop.test(sum(matches_LL_vs_DE$LLW == 1), nrow(matches_LL_vs_DE), conf.level = 0.95)

# LL vs SE

matches_LL_vs_SE <- datos_combinados %>%
  filter(LL == "Yes" & (winner_entry == "SE" | loser_entry == "SE"))


prop.test(sum(matches_LL_vs_SE$LLW == 1), nrow(matches_LL_vs_SE), conf.level = 0.95)

# LL vs Alt

matches_LL_vs_alt <- datos_combinados[datos_combinados$LL == "Yes" & 
                                     (datos_combinados$winner_entry %in% c("Alt", "ALT") | 
                                      datos_combinados$loser_entry %in% c("Alt", "ALT")), ]


# prop.test(sum(matches_LL_vs_alt$LLW == 1), nrow(matches_LL_vs_alt), conf.level = 0.95)
## Not found


# LL vs normal entry

matches_LL_vs_normal <- datos_combinados[
  datos_combinados$LL == "Yes" & 
  (is.na(datos_combinados$winner_entry) | datos_combinados$winner_entry == "" | 
   is.na(datos_combinados$loser_entry) | datos_combinados$loser_entry == ""), ]


prop.test(sum(matches_LL_vs_normal$LLW == 1), nrow(matches_LL_vs_normal), conf.level = 0.95)
## Not found



```


```{r}

# PR vs LL

matches_PR_LL <- datos_combinados %>%
  filter(
    PR == "Yes" &  
    (winner_entry == "LL" | loser_entry == "LL"))

prop.test(sum(matches_PR_LL$PRW == 1), nrow(matches_PR_LL), conf.level = 0.95)


# PR vs PR

matches_PR_vs_PR <- datos_combinados %>% filter(
    LL == "Yes" &  (winner_entry == "PR" & loser_entry == "PR"))
#prop.test(sum(matches_PR_vs_PR$PRW == 1), nrow(matches_PR_vs_PR), conf.level = 0.95)
#0/0

# PR vs Direct entry

matches_PR_vs_DE <- datos_combinados[datos_combinados$PR == "Yes" & 
                                     (datos_combinados$winner_entry %in% c("Q", "WC", "ITF") | 
                                      datos_combinados$loser_entry %in% c("Q", "WC", "ITF")), ]


prop.test(sum(matches_PR_vs_DE$PRW == 1), nrow(matches_PR_vs_DE), conf.level = 0.95)


# PR vs SE

matches_PR_vs_SE <- datos_combinados %>%
  filter(PR == "Yes" & (winner_entry == "SE" | loser_entry == "SE"))


prop.test(sum(matches_PR_vs_SE$PRW == 1), nrow(matches_PR_vs_SE), conf.level = 0.95)

# PR vs Alt

matches_PR_vs_alt <- datos_combinados[datos_combinados$PR == "Yes" & 
                                     (datos_combinados$winner_entry %in% c("Alt", "ALT") | 
                                      datos_combinados$loser_entry %in% c("Alt", "ALT")), ]


prop.test(sum(matches_PR_vs_alt$PRW == 1), nrow(matches_PR_vs_alt), conf.level = 0.95)



# PR vs normal entry

matches_PR_vs_normal <- datos_combinados[datos_combinados$PR == "Yes" & 
                                           (is.na(datos_combinados$winner_entry) | datos_combinados$winner_entry == "" | 
   is.na(datos_combinados$loser_entry) | datos_combinados$loser_entry == ""), ]

prop.test(sum(matches_PR_vs_normal$PRW == 1), nrow(matches_PR_vs_normal), conf.level = 0.95)
## Not found

```






