---
title: "Models"
author: "Ana Abad"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
load("ATP_final.RData")
data <- datos_PR_periodo

library(car)
library(dplyr)
library(DHARMa)
library(corrplot)
library(ggplot2)
library(dplyr)
library(data.table)
library(mgcv)
library(mgcViz)
library(lme4)
library(MASS)
library(splines)
library(caret)
library(broom)
library(emmeans)


```


Firstly, we determine which variables we want to include in the model: surface, round, tourney_level, year, loser_hand, winner_hand, dif_age, dif_rank
Response: PRW and PRL



```{r}
# Factors / qualitative 
data$tourney_level    <- factor(data$tourney_level, ordered = F)
data$surface          <- as.factor(data$surface)


data <- data %>%
  mutate(round = case_when(
    grepl("^R[0-9]+$", round) ~ "Preliminary", 
    round %in% c("RR", "BR") ~ "Preliminary",  
    TRUE ~ round
  ))

data <- data %>%
  mutate(round = case_when(
    round %in% c("Q1", "Q2", "Q3") ~ "Quarters",  
    TRUE ~ round
  ))
data$round      <- factor(data$round, ordered = F)



data$best_of          <- factor(data$best_of)

# Response variables
data$PRW              <- as.factor(data$PRW)
data$PRL              <- as.factor(data$PRL)

# Numeric
data$year             <- as.numeric(data$year)
data$dif_age          <- as.numeric(data$dif_age)
data$dif_rank         <- as.numeric(data$dif_rank)


#### Reference level
data$tourney_level <- relevel(data$tourney_level, ref = "Masters") # Masters
data$surface          <- relevel(data$surface,          ref = "Grass")
data$round      <- relevel(data$round,      ref = "Final") #Final

```


```{r}
data_PR <- subset(data, PR == "Yes")
```

# Independence 



```{r}


fisher.test(data$surface, data$round, simulate.p.value=TRUE)
chisq.test(data$surface, data$tourney_level)
fisher.test(data$surface, data$year, simulate.p.value=TRUE)
chisq.test(data$surface, data$loser_hand)
chisq.test(data$surface, data$winner_hand)


fisher.test(data$round, data$tourney_level, simulate.p.value=TRUE)
fisher.test(data$round, data$year, simulate.p.value=TRUE)
fisher.test(data$round, data$loser_hand, simulate.p.value=TRUE)
fisher.test(data$round, data$winner_hand, simulate.p.value=TRUE)


fisher.test(data$tourney_level, data$year, simulate.p.value=TRUE)
chisq.test(data$tourney_level, data$loser_hand)
chisq.test(data$tourney_level, data$winner_hand)


fisher.test(data$year, data$loser_hand, simulate.p.value=TRUE)
fisher.test(data$year, data$winner_hand, simulate.p.value=TRUE)


chisq.test(data$loser_hand, data$winner_hand)


fisher.test(data$best_of, data$surface, simulate.p.value=TRUE)
fisher.test(data$best_of, data$round, , simulate.p.value=TRUE)
fisher.test(data$best_of, data$tourney_level, , simulate.p.value=TRUE)
chisq.test(data$best_of, data$winner_hand)
fisher.test(data$best_of, data$loser_hand, , simulate.p.value=TRUE)
fisher.test(data$best_of, data$year, , simulate.p.value=TRUE)


```


# Interaction

```{r}

cor_data <- data[, c("year", "dif_rank", "dif_age")]
cor_matrix <- cor(cor_data, use = "complete.obs")
corrplot(cor_matrix, method = "shade", type = "lower")

```

```{r}

ggplot(data, aes(x = surface, y = dif_age, fill = round)) +
  geom_boxplot() +
  labs(title = "Interaction: Surface, Round, and dif_age") +
  theme_minimal()
vcd::assocstats(table(data$surface, data$round))
```
Interaction between surface and round is low

```{r}

vcd::assocstats(table(data$year, data$round))
ggplot(data, aes(x = year, fill = round)) +
  geom_bar(position = "dodge") +
  labs(title = "Association Between Year and Round", x = "year", y = "Count") +
  theme_minimal()

vcd::assocstats(table(data$year, data$surface))
ggplot(data, aes(x = year, fill = surface)) +
  geom_bar(position = "dodge") +
  labs(title = "Association Between Surface and Year", x = "year", y = "Count") +
  theme_minimal()

vcd::assocstats(table(data$year, data$winner_hand))
ggplot(data, aes(x = year, fill = winner_hand)) +
  geom_bar(position = "dodge") +
  labs(title = "Association Between year and winner_hand", x = "Surface", y = "Count") +
  theme_minimal()

vcd::assocstats(table(data$year, data$loser_hand))
ggplot(data, aes(x = year, fill = loser_hand)) +
  geom_bar(position = "dodge") +
  labs(title = "Association Between year and loser_hand", x = "year", y = "Count") +
  theme_minimal()
```

There seems to be interaction between year and all categorical variables, but the associations are low. 

```{r}
vcd::assocstats(table(data$surface, data$round))
ggplot(data, aes(x = surface, fill = round)) +
  geom_bar(position = "dodge") +
  labs(title = "Association Between Surface and Round", x = "Surface", y = "Count") +
  theme_minimal()

vcd::assocstats(table(data$surface, data$winner_hand))
ggplot(data, aes(x = surface, fill = winner_hand)) +
  geom_bar(position = "dodge") +
  labs(title = "Association Between Surface and winner_hand", x = "Surface", y = "Count") +
  theme_minimal()


vcd::assocstats(table(data$surface, data$loser_hand))
ggplot(data, aes(x = surface, fill = loser_hand)) +
  geom_bar(position = "dodge") +
  labs(title = "Association Between Surface and loser_hand", x = "Surface", y = "Count") +
  theme_minimal()


vcd::assocstats(table(data$loser_hand, data$winner_hand))
ggplot(data, aes(x = loser_hand, fill = winner_hand)) +
  geom_bar(position = "dodge") +
  labs(title = "Association Between loser_hand and winner_hand", x = "loser_hand", y = "Count") +
  theme_minimal()
```
The associations and the correlations are low.

```{r}
vcd::assocstats(table(data$best_of, data$winner_hand))
ggplot(data, aes(x = best_of, fill = winner_hand)) +
  geom_bar(position = "dodge") +
  labs(title = "Association Between best_of and winner_hand", x = "best_of", y = "Count") +
  theme_minimal()

vcd::assocstats(table(data$best_of, data$loser_hand))
ggplot(data, aes(x = best_of, fill = loser_hand)) +
  geom_bar(position = "dodge") +
  labs(title = "Association Between best_of and Loser_hand", x = "best_of", y = "Count") +
  theme_minimal()

vcd::assocstats(table(data$best_of, data$surface))
ggplot(data, aes(x = best_of, fill = surface)) +
  geom_bar(position = "dodge") +
  labs(title = "Association Between best_of and surface", x = "best_of", y = "Count") +
  theme_minimal()

vcd::assocstats(table(data$best_of, data$tourney_level))
ggplot(data, aes(x = best_of, fill = tourney_level)) +
  geom_bar(position = "dodge") +
  labs(title = "Association Between best_of and tourney_level", x = "best_of", y = "Count") +
  theme_minimal()

vcd::assocstats(table(data$best_of, data$round))
ggplot(data, aes(x = best_of, fill = round)) +
  geom_bar(position = "dodge") +
  labs(title = "Association Between best_of and round", x = "best_of", y = "Count") +
  theme_minimal()

```
The associations and the correlations are low, except for tourney_level, which was high, and round, which were moderate. Therefore, best_of will be excluded from the possible variables to avoid multicolinearity. 


# Lineality

The + 0.5 is a continuity correction to avoid division by zero if a count is zero.

## Years

```{r}

table_year <- table(data_PR$year, data_PR$LLW) 
log_odd_year <- log((table_year[,2]+.5) / (table_year[,1]+.5)) 

year <- as.numeric(rownames(table_year)) 

plot(year,log_odd_year,xlab="Years",ylab="Log-odds",las=1, main="LLW")
lines(lowess(log_odd_year~year), col=4, lwd=2)


##

table_year <- table(data_PR$year, data_PR$LLL) 
log_odd_year <- log((table_year[,2]+.5) / (table_year[,1]+.5)) 

year <- as.numeric(rownames(table_year)) 

plot(year,log_odd_year,xlab="Years",ylab="Log-odds",las=1, main="LLL")
lines(lowess(log_odd_year~year), col=4, lwd=2)


##

table_year <- table(data_PR$year, data_PR$PRW) 
log_odd_year <- log((table_year[,2]+.5) / (table_year[,1]+.5)) 

year <- as.numeric(rownames(table_year)) 

plot(year,log_odd_year,xlab="Years",ylab="Log-odds",las=1, main="PRW")
lines(lowess(log_odd_year~year), col=4, lwd=2)

##

table_year <- table(data_PR$year, data_PR$PRL) 
log_odd_year <- log((table_year[,2]+.5) / (table_year[,1]+.5)) 

year <- as.numeric(rownames(table_year)) 

plot(year,log_odd_year,xlab="Years",ylab="Log-odds",las=1, main="PRL")
lines(lowess(log_odd_year~year), col=4, lwd=2)


```


## dif_age 

```{r}

table_dif_age <- table(data_PR$dif_age, data_PR$LLW) 
log_odd_dif_age <- log((table_dif_age[,2]+.5) / (table_dif_age[,1]+.5)) 

dif_age <- as.numeric(rownames(table_dif_age)) 

plot(dif_age,log_odd_dif_age,xlab="dif_ages",ylab="Log-odds",las=1, main="LLW")
lines(lowess(log_odd_dif_age~dif_age), col=4, lwd=2)


##

table_dif_age <- table(data_PR$dif_age, data_PR$LLL) 
log_odd_dif_age <- log((table_dif_age[,2]+.5) / (table_dif_age[,1]+.5)) 

dif_age <- as.numeric(rownames(table_dif_age)) 

plot(dif_age,log_odd_dif_age,xlab="dif_ages",ylab="Log-odds",las=1, main="LLL")
lines(lowess(log_odd_dif_age~dif_age), col=4, lwd=2)


##

table_dif_age <- table(data_PR$dif_age, data_PR$PRW) 
log_odd_dif_age <- log((table_dif_age[,2]+.5) / (table_dif_age[,1]+.5)) 

dif_age <- as.numeric(rownames(table_dif_age)) 

plot(dif_age,log_odd_dif_age,xlab="dif_ages",ylab="Log-odds",las=1, main="PRW")
lines(lowess(log_odd_dif_age~dif_age), col=4, lwd=2)

##

table_dif_age <- table(data_PR$dif_age, data_PR$PRL) 
log_odd_dif_age <- log((table_dif_age[,2]+.5) / (table_dif_age[,1]+.5)) 

dif_age <- as.numeric(rownames(table_dif_age)) 

plot(dif_age,log_odd_dif_age,xlab="dif_ages",ylab="Log-odds",las=1, main="PRL")
lines(lowess(log_odd_dif_age~dif_age), col=4, lwd=2)


```

## dif_rank

```{r}

table_dif_rank <- table(data_PR$dif_rank, data_PR$LLW) 
log_odd_dif_rank <- log((table_dif_rank[,2]+.5) / (table_dif_rank[,1]+.5)) 

dif_rank <- as.numeric(rownames(table_dif_rank)) 

plot(dif_rank,log_odd_dif_rank,xlab="dif_ranks",ylab="Log-odds",las=1, main="LLW")
lines(lowess(log_odd_dif_rank~dif_rank), col=4, lwd=2)


##

table_dif_rank <- table(data_PR$dif_rank, data_PR$LLL) 
log_odd_dif_rank <- log((table_dif_rank[,2]+.5) / (table_dif_rank[,1]+.5)) 

dif_rank <- as.numeric(rownames(table_dif_rank)) 

plot(dif_rank,log_odd_dif_rank,xlab="dif_ranks",ylab="Log-odds",las=1, main="LLL")
lines(lowess(log_odd_dif_rank~dif_rank), col=4, lwd=2)


##

table_dif_rank <- table(data_PR$dif_rank, data_PR$PRW) 
log_odd_dif_rank <- log((table_dif_rank[,2]+.5) / (table_dif_rank[,1]+.5)) 

dif_rank <- as.numeric(rownames(table_dif_rank)) 

plot(dif_rank,log_odd_dif_rank,xlab="dif_ranks",ylab="Log-odds",las=1, main="PRW")
lines(lowess(log_odd_dif_rank~dif_rank), col=4, lwd=2)

##

table_dif_rank <- table(data_PR$dif_rank, data_PR$PRL) 
log_odd_dif_rank <- log((table_dif_rank[,2]+.5) / (table_dif_rank[,1]+.5)) 

dif_rank <- as.numeric(rownames(table_dif_rank)) 

plot(dif_rank,log_odd_dif_rank,xlab="dif_ranks",ylab="Log-odds",las=1, main="PRL")
lines(lowess(log_odd_dif_rank~dif_rank), col=4, lwd=2)


```

### Conclusions: 

Most of the variables are categorical, and therefore the concept of linearity does not directly apply to them. Years variable looks linear for lucky losers, but has two very differenciated slopes after 2005 for PR. dif_age and dig_rank show similar graphics (V-shapped) for both categories, clearly non linear. 


# PR

# Models

Despite non-linearity, a glm will be fitted although it could not be the best approach. 


```{r}



data_grouped_PRW <- data_PR[,c("tourney_level", "year", "surface", "winner_hand", "loser_hand", "dif_age", "dif_rank", "games", "PRW", "loser_id", "winner_id", "winner_age", "loser_age", "winner_rank", "loser_rank")]


data_grouped_PRW <- data_grouped_PRW[complete.cases(data_grouped_PRW) & data_grouped_PRW$games!=0,]

# Only includes completed data and with a number of games superior to 0.

#### Models

glm_1_PR <- glm(PRW ~ tourney_level + year + surface + 
                    dif_age + dif_rank + winner_hand + loser_hand,
                    family = "binomial", offset = log(games), data = data_grouped_PRW)

summary(glm_1_PR)



testDispersion(glm_1_PR)

## Pearson dispersion statistic
sum(residuals(glm_1_PR, type = "pearson")^2) / glm_1_PR$df.residual




## Selection by AIC criteria using step function

step_1_PR <- step(glm_1_PR, criterion = "BIC")
summary(step_1_PR)

testDispersion(step_1_PR)

## Pearson dispersion statistic
sum(residuals(step_1_PR, type = "pearson")^2) / step_1_PR$df.residual





## Avoiding variables with problems with linearity 


glm_2_PR <- glm(PRW ~ tourney_level + year + surface + 
                       winner_hand + loser_hand,
                    family = "binomial", offset = log(games), data = data_grouped_PRW)

summary(glm_2_PR)


testDispersion(glm_2_PR)

## Pearson dispersion statistic
sum(residuals(glm_2_PR, type = "pearson")^2) / glm_2_PR$df.residual




## Selection by AIC criteria using step function

step_2_PR <- step(glm_2_PR, criterion = "BIC")
summary(step_2_PR)

testDispersion(step_2_PR)

## Pearson dispersion statistic
sum(residuals(step_2_PR, type = "pearson")^2) / step_2_PR$df.residual


# Retested with interactions between surface and tourney_level and winner/loser_hand interaction


glm_3_PR <- glm(PRW ~ tourney_level + year + surface + 
                       dif_age + dif_rank + winner_hand + loser_hand + winner_hand * loser_hand + tourney_level * surface, 
                   family = "binomial", offset = log(games), data = data_grouped_PRW) 

summary(glm_3_PR)


testDispersion(glm_3_PR)

## Pearson dispersion statistic
sum(residuals(glm_3_PR, type = "pearson")^2) / glm_3_PR$df.residual



step_3_PR <- step(glm_3_PR, criterion = "BIC")
summary(step_3_PR)
vif(step_3_PR)

testDispersion(step_3_PR)

## Pearson dispersion statistic
sum(residuals(step_3_PR, type = "pearson")^2) / step_3_PR$df.residual




```








## Overdispersion


```{r}


# The best model at the moment is step_3 (lowest AIC)


testDispersion(step_3_PR)

## Pearson dispersion statistic
sum(residuals(step_3_PR, type = "pearson")^2) / step_3_PR$df.residual
```

There seems to be overdispersion (Pearson > 1) 

### Transformations

```{r}

glm_4_PR <- glm(PRW ~ tourney_level + year + surface + 
                       log(abs(dif_rank)) + dif_age+ winner_hand + loser_hand + winner_hand * loser_hand + tourney_level * surface, 
                   family = "binomial", offset = log(games), data = data_grouped_PRW) 

summary(glm_4_PR)
testDispersion(glm_4_PR)

## Pearson dispersion statistic
sum(residuals(glm_4_PR, type = "pearson")^2) / glm_4_PR$df.residual




step_4_PR <- step(glm_4_PR, criterion = "BIC")
summary(step_4_PR)
sum(residuals(step_4_PR, type = "pearson")^2) / step_4_PR$df.residual
testDispersion(step_4_PR)

```
The AIC is worse, but the coefficient is better. 


```{r}

glm_5_PR <- glm(PRW ~ tourney_level + year + surface + 
                       tan(dif_rank+1900) + tan(dif_age+20) + winner_hand + loser_hand + winner_hand * loser_hand + tourney_level * surface, 
                   family = "binomial", offset = log(games), data = data_grouped_PRW) 

summary(glm_5_PR)
testDispersion(glm_5_PR)

## Pearson dispersion statistic
sum(residuals(glm_5_PR, type = "pearson")^2) / glm_5_PR$df.residual




step_5_PR <- step(glm_5_PR, criterion = "BIC")
summary(step_5_PR)
sum(residuals(step_5_PR, type = "pearson")^2) / step_5_PR$df.residual



testDispersion(step_5_PR)

```


```{r}
plot((dif_rank))
```


### Converting dif_rank and dif_age to categorical: category not used 

```{r}

new_data_grouped_PRW <- data_grouped_PRW

# Define the categories based on dif_rank- WIB: Winner is better, LIB: Loser is better 
new_data_grouped_PRW$dif_rank_category <- cut(
  new_data_grouped_PRW$dif_rank,
  breaks = c(-Inf, -1200, -900, -600, -300, 0, 300, 600, 900, 1200, Inf),
  labels = c(
    "Very High WIB",  # Below -1200
    "High WIB",       # -1200 to -900
    "Moderate WIB",   # -900 to -600
    "Low WIB",        # -600 to -300
    "Very Low WIB",   # -300 to 0
    "Very Low LIB",   # 0 to 300
    "Low LIB",        # 300 to 600
    "Moderate LIB",   # 600 to 900
    "High LIB",       # 900 to 1200
    "Very High LIB"   # More than 1200
  ),
  include.lowest = TRUE,
  right = FALSE 
)

new_data_grouped_PRW$dif_age_category <- cut(
  new_data_grouped_PRW$dif_age,
  breaks = c(-Inf, -15, -10, -5, 0, 5, 10, 15, Inf),
  labels = c(
    "Very High WIB",  # Below -15
    "High WIB",       # -15 to -10
    "Moderate WIB",   # -10 to -5
    "Low WIB",        # -5 to 0
    "Very Low WIB",   # 0 to 5
    "Very Low LIB",   # 5 to 10
    "Low LIB",        # 10 to 15
    "High LIB"        # Above 15
  ),
  include.lowest = TRUE,
  right = FALSE  
)




```

```{r}

glm_6_PR <- glm(PRW ~ tourney_level + year + surface + 
                       dif_rank_category + dif_age_category + winner_hand + loser_hand + winner_hand * loser_hand + tourney_level * surface, 
                   family = "binomial", offset = log(games), data = new_data_grouped_PRW) 

summary(glm_6_PR)
step_6_PR <- step(glm_6_PR, criterion = "BIC")
summary(step_6_PR)
sum(residuals(step_6_PR, type = "pearson")^2) / step_6_PR$df.residual



testDispersion(step_6_PR)
```

Using a more "restrictive" categorization, based on https://data.scorenetwork.org/tennis/australian-open.html

```{r}



new_data_grouped_PRW_2 <- data_grouped_PRW %>%
  mutate(
    rank_category = case_when(
      abs(dif_rank) <= 30 ~ "Low", 
      abs(dif_rank) > 30 & abs(dif_rank) <= 75 ~ "Medium",  
      abs(dif_rank) > 75 ~ "High"    
    ),
    age_category = case_when(
      abs(dif_age) <= 5 ~ "Low",  
      abs(dif_age) > 5 & abs(dif_age) <= 15 ~ "Medium", 
      abs(dif_age) > 15 ~ "High"    
    )
  )

glm_7_PR <- glm(PRW ~ tourney_level + year + surface + 
                       rank_category + winner_hand + loser_hand + 
                  winner_hand * loser_hand + tourney_level * surface + age_category, 
                   family = "binomial", offset = log(games), 
                data = new_data_grouped_PRW_2) 

summary(glm_7_PR)

testDispersion(glm_7_PR)

## Pearson dispersion statistic
sum(residuals(glm_7_PR, type = "pearson")^2) / glm_7_PR$df.residual



step_7_PR <- step(glm_7_PR, criterion = "BIC")
summary(step_7_PR)
sum(residuals(step_7_PR, type = "pearson")^2) / step_7_PR$df.residual



testDispersion(step_7_PR)

```

Using a different categorization


```{r}


new_data_grouped_PRW_3 <- data_grouped_PRW %>%
  mutate(
    rank_category = case_when(
      dif_rank <= -30 ~ "Winnerishigherranked", 
      dif_rank > -30 & dif_rank <= 30 ~ "Medium",  
      dif_rank > 30 ~ "Loserishigherranked"    
    ),
    age_category = case_when(
      dif_age <= -5 ~ "YoungerWinner",  
      dif_age > -5 & dif_age <= 5 ~ "Neutral", 
      dif_age > 5 ~ "OlderWinner"    
    )
  )
new_data_grouped_PRW_3$rank_category <- factor(new_data_grouped_PRW_3$rank_category)
new_data_grouped_PRW_3$rank_category  <- relevel(new_data_grouped_PRW_3$rank_category, ref = "Medium")


glm_8_PR <- glm(PRW ~ tourney_level + year + surface + 
                       rank_category + winner_hand + loser_hand + winner_hand * loser_hand + tourney_level * surface + age_category, 
                   family = "binomial", offset = log(games), data = new_data_grouped_PRW_3) 

summary(glm_8_PR)
testDispersion(glm_8_PR)

## Pearson dispersion statistic
sum(residuals(glm_8_PR, type = "pearson")^2) / glm_8_PR$df.residual





step_8_PR <- step(glm_8_PR, criterion = "BIC")
summary(step_8_PR)
sum(residuals(step_8_PR, type = "pearson")^2) / step_8_PR$df.residual



testDispersion(step_8_PR)

```




### GAM

```{r}



# Fitting GAM with smooth terms for dif_rank and dif_age
gam_model_PR <- gam(PRW ~ s(dif_rank, bs = "cr") + s(dif_age, bs = "cr") + #cubic regresion
                   tourney_level + surface + winner_hand + loser_hand,
                 family = binomial, 
                 data = data_grouped_PRW)


summary(gam_model_PR)


sum(residuals(gam_model_PR, type = "pearson")^2) / gam_model_PR$df.residual
testDispersion(gam_model_PR)




#Since there is no step.gam function, only the significant variables will be chosen

gam_model_PR_2 <- gam(PRW ~ s(dif_rank, bs = "cr") + s(dif_age, bs = "cr"),
                 family = binomial, 
                 data = data_grouped_PRW)
summary(gam_model_PR_2)
AIC(gam_model_PR)
AIC(gam_model_PR_2)


sum(residuals(gam_model_PR_2, type = "pearson")^2) / gam_model_PR_2$df.residual
testDispersion(gam_model_PR_2)

```
## Natural splines 

```{r}



splines_PR_1 <- glm(PRW ~ ns(dif_age, 3) + ns(dif_rank, 3) + tourney_level + surface + winner_hand + loser_hand,
                 family = binomial, 
                 data = new_data_grouped_PRW)

summary(splines_PR_1)


testDispersion(splines_PR_1)

## Pearson dispersion statistic
sum(residuals(splines_PR_1, type = "pearson")^2) / splines_PR_1$df.residual






step_PR_splines <- step(splines_PR_1, criterion = "BIC")


sum(residuals(step_PR_splines, type = "pearson")^2) / step_PR_splines$df.residual
testDispersion(step_PR_splines)


splines_PR_2 <- glm(PRW ~ ns(dif_age, 2) + ns(dif_rank, 2) + tourney_level + surface + winner_hand + loser_hand,
                 family = binomial, 
                 data = new_data_grouped_PRW)

summary(splines_PR_2)

sum(residuals(splines_PR_2, type = "pearson")^2) / splines_PR_2$df.residual
testDispersion(splines_PR_2)



step_PR_splines_2 <- step(splines_PR_2, criterion = "BIC")
sum(residuals(step_PR_splines_2, type = "pearson")^2) / step_PR_splines_2$df.residual
testDispersion(step_PR_splines_2)


```

```{r}

glm_9_PR <- glm(PRW ~ tourney_level + year + surface + games +
                    dif_age + dif_rank + winner_hand + loser_hand,
                    family = "binomial",  data = data_grouped_PRW)

summary(glm_9_PR)



testDispersion(glm_9_PR)

## Pearson dispersion statistic
sum(residuals(glm_9_PR, type = "pearson")^2) / glm_9_PR$df.residual





step_9_PR <- step(glm_9_PR, criterion = "BIC")
summary(step_9_PR)

testDispersion(step_9_PR)

## Pearson dispersion statistic
sum(residuals(step_9_PR, type = "pearson")^2) / step_9_PR$df.residual
```

#### Model 10: no dif, offset

```{r}

glm_10_PR <- glm(PRW ~ tourney_level + year + surface + 
                    loser_age + winner_age + winner_rank + loser_rank + winner_hand + loser_hand,
                    family = "binomial", offset = log(games),  data = data_grouped_PRW)

summary(glm_10_PR)



testDispersion(glm_10_PR)

## Pearson dispersion statistic
sum(residuals(glm_10_PR, type = "pearson")^2) / glm_10_PR$df.residual




## Selection by AIC criteria using step function

step_10_PR <- step(glm_10_PR, criterion = "BIC")
summary(step_10_PR)

testDispersion(step_10_PR)

## Pearson dispersion statistic
sum(residuals(step_10_PR, type = "pearson")^2) / step_10_PR$df.residual
```



## Best model PR

```{r}



summary(step_8_PR)
plot(step_8_PR)
vif(step_8_PR)
testDispersion(step_8_PR)
sum(residuals(step_8_PR, type = "pearson")^2) / step_8_PR$df.residual

exp(cbind(OR = coef(step_8_PR), confint(step_8_PR)))


```

# Validation

```{r}




incomplete_data_PR <- data_PR[!complete.cases(data_PR) & data_PR$games != 0, ]
incomplete_data_PR <- incomplete_data_PR %>%
  mutate(
    rank_category = case_when(
      dif_rank <= -30 ~ "Winnerishigherranked", 
      dif_rank > -30 & dif_rank <= 30 ~ "Medium",  
      dif_rank > 30 ~ "Loserishigherranked"    
    ),
    age_category = case_when(
      dif_age <= -5 ~ "YoungerWinner",  
      dif_age > -5 & dif_age <= 5 ~ "Neutral", 
      dif_age > 5 ~ "OlderWinner"    
    )
  )
incomplete_data_PR$surface[!incomplete_data_PR$surface %in% levels(incomplete_data_PR$surface)] <- NA

complete_data_PR_validation <- incomplete_data_PR[complete.cases(incomplete_data_PR[, c("rank_category", "age_category")]), ]



```


## PR

```{r}

pred_probs <- predict(step_8_PR, newdata = complete_data_PR_validation, type = "response")
pred_classes <- ifelse(pred_probs > 0.5, "1", "0")

pred_classes <- factor(pred_classes, levels = c("0", "1"))
complete_data_PR_validation$PRW <- factor(complete_data_PR_validation$PRW, levels = c("0", "1"))


confusionMatrix(pred_classes, complete_data_PR_validation$PRW)


```

# Forest plot


```{r}
model_summary_2 <- tidy(step_8_PR, conf.int = TRUE, exp = TRUE)


b <- ggplot(model_summary_2, aes(y = term, x = estimate, xmin = conf.low, xmax = conf.high)) +
  geom_pointrange() +
  geom_vline(xintercept = 1, linetype = "dashed", color = "darkblue") +
  theme_minimal() +
  labs(x = "Odds Ratio", y = "Predictor", title = "Protected Ranking model") +
  theme(axis.text.y = element_text(size = 12), axis.title = element_text(size = 14))


ggsave("PRforestplot.tiff", plot = b, device = "tiff", width = 8, height = 6, dpi = 300)


```


# emmeans


```{r}

# PR


emm_age <- emmeans(step_8_PR, ~age_category, type = 'response')
emm_rank <- emmeans(step_8_PR, ~rank_category, type = 'response')
emm_level <- emmeans(step_8_PR, ~tourney_level, type = 'response')


summary(emm_age)
summary(emm_rank)
summary(emm_level)

```


